---
title: "P8105_hw_yx2510"
author: "Yi Xiao"
date: "28/09/2018"
output: html_document
---
load package
```{r}
library(tidyverse)
```

Problem 1
### 1. read and clean the data
```{r}
nyc_transit_data =
        read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
        janitor::clean_names() %>%
        select(line:entry,vending, ada)

nyc_transit_data$entry = ifelse(nyc_transit_data$entry == "yes", TRUE, FALSE)

```
After importing the file using read_csv() function, I found that there are upper cases and spaces in the variable names, so I first use the clean_name() for the name cleaning. Then I used the select() function to include those variables we are interested in. At last, ifelse() was used to convert the entry status to corresponding logical variables. There were `r ncol(nyc_transit_data)` columns and `r nrow(nyc_transit_data)`. However, the dataset is not clean since the route information was scattered across several colunms. Hence, data cleaning is required to gather the information in a single column.

```{r}

```


